---
title: "ST558 - Project 2 - Predictive Modeling"
author: "Jasmine Wang & Ilana Feldman"
date: "10/31/2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction - done by Jasmine

briefly describes the data 
briefly describes the variables you have to work with (describe what you want to use)

purpose of the analysis
methods you will use to model the response (more details in modeling section)

61 variables (only 58 predictive variables, 2 non-predictive), target response is "shares".

# Data 

```{r eval=TRUE}
library(tidyverse)
library(knitr)
library(caret)

allnews <- read_csv("C:/Users/peach/Documents/ST558/ST558_repos/ST558-Project-2/_Data/OnlineNewsPopularity.csv", 
                 col_names = TRUE)
dim(allnews)
all_news <- allnews %>% mutate(class_shares = if_else(shares < 1400, 0, 1))
news <- all_news %>% filter(data_channel_is_lifestyle == 1) %>% select(
  -data_channel_is_lifestyle, -data_channel_is_entertainment, -data_channel_is_bus, -data_channel_is_socmed, 
  -data_channel_is_tech, -data_channel_is_world, -url, -timedelta)
#news <- allnews %>% filter(data_channel_is_lifestyle == 1) %>% select(
#  n_tokens_title, n_tokens_content, num_hrefs, num_imgs, num_videos, kw_avg_min, kw_avg_max, avg_negative_polarity, 
#  avg_positive_polarity, title_subjectivity, title_sentiment_polarity, shares)
news

set.seed(388588)
sharesIndex <- createDataPartition(news$shares, p = 0.7, list = FALSE)
train <- news[sharesIndex, ]
test <- news[-sharesIndex, ]

train1 <- train %>% select(-class_shares)
test1 <- test %>% select(-class_shares)
fit1 <- lm(shares ~ . , data = train1)
summary(fit1)

cv_fit1 <- train(shares ~ . , 
                 data=train1,
                 method = "lm",
                 preProcess = c("center", "scale"),
                 trControl = trainControl(method = "cv", number = 10))
cv_fit1


```


```{r eval=TRUE}

<- train %>% select(-shares)
<- test %>% select(-shares)




```
# Summarizations

## Numerical Summaries

## Visualizations

# Modeling

## Linear Regression (2)

one each

## Random Forest 

Ilana

## Boosted Tree

Jasmine

# Comparison

# Automation


You can also embed plots, for example:

```{r echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
